Wireless Automobile Detection, License Plate Processing, and Data Availability Network

### 2014-04-30 paper/paper.tex
Stuff

\section{Related Work}
{\color{red}\textbf{NOT COMPLETE}}
The ACMx group at CSM has designed a system named SmartLots to track the number of vehicles in a parking lot. Their system is a wireless sensor network consisting of sensor nodes and a central base station. The sensor nodes are based on the Arduino Fio microcontroller platform equipped with triaxial magnetometers and XBee Pro radios. The base station will be a Raspberry Pi connected to the internet via Ethernet. The information collected by the sensor nodes is transmitted to the base station and forwarded to a server running a small web application. The system is described in \cite{stillwell2013} and \cite{parkingWiki}. \cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using the Arduino Fio and the Raspberry Pi. This system uses the IEEE 802.15.4 communications standard to communicate automobile detection data from the sensor nodes to the base station. \cite{parkingWiki} describes the ongoing status of the ACMx project. At the time of writing this proposal, the ACMx team has created a website that will display the status of the CTLM upper and lower lots once the monitoring system has been deployed. \cite{parkingWiki} is updated periodically with new information about project progress. The project described in this proposal will augment and extend the ACMx project while collaborating closely with Stillwell (ACMx president and coauthor of \cite{stillwell2013}), with the goal of a real world deployment by May 2014.

\section{Project Overview}
{\color{red}\textbf{NOT COMPLETE}}
\subsection{Overview}
The project we are proposing will be an extension of the ACMx SmartLots project. We will collaborate closely with Stillwell and the ACMx group to create and deploy the first working parking lot monitoring system on the CSM campus. Our system will consist of several components, each of which will be either an extension of a current SmartLots system or an entirely new subsystem that will be integrated into the existing system. Figure \ref{fig:system} shows the overall architecture of our system.

\begin{figure}
\begin{center}
\includegraphics[width=4.5in]{architecture}
\end{center}
\caption{Overall System Architecture}
\label{fig:system}
\end{figure}

The system will rely on three main components:
\begin{enumerate}
\item Detection Subsystem,
\item Central Base Station,
\item and Server.
\end{enumerate}
The detection subsystem will be made up of a small number of sensor nodes, placed strategically at parking lot entrances. Each sensor node will contain an Arduino Fio equipped with a triaxial magnetometer that will be used to detect when a car passes the sensor. The work to determine the hardware configuration and software for the Arduino Fios has already been done by the ACMx group. Our contribution will be to also equip each node with a Raspberry Pi mini-computer and camera, which will take pictures of the license plates of cars as they pass by. When a car passes a sensor node, the magnetometer will detect the car, and the Arduino Fio will send an interrupt to the Raspberry Pi via the XBee radio, waking it up so it can take a picture of the license plate.

The central base station will be a Raspberry Pi that will receive all the data from the sensor nodes via transmission from the XBee Pro radios attached to the sensor nodes. The base station Raspberry Pi will be centrally located in a nearby building with internet connectivity, and will have a camera attached that will take periodic images of the lots from above. The base station will use its network connection to forward the data to the ACMx server.

The server will be running a web application with an HTTP interface that will both receive the data from the base station and route it to the correct part of the application (database, image processor, etc.). This interface will be the method of communication between the server and the base station. When data is received from the base station, the application will process it, store it in a database, and display it appropriately on the website. The following sections describe the architecture and function of each of the aforementioned subsystems in detail. The detection subsystem is broken up into two parts: Automobile Detection (section 2.2) and License Plate Image Acquisition (section 2.3). The server subsystem is also broken into two parts: Server Processing (section 2.5) and Web Application (section 2.6).

\subsection{Automobile Detection}
{\color{red}\textbf{NOT COMPLETE}}
The automobile detection subsystem will provide a means for detecting ingress and egress of vehicles from a parking lot. This subsystem is to be placed on the side of the road next to each parking lot entrance and exit. Automobiles will be detected using a magnetometer which perceives the induced change in the local magnetic field as the metallic structure of the vehicle passes by, as described in \cite{stillwell2013}. The automobile detection subsystem will be based around the commercially available Arduino Fio 16-bit  platform which utilizes the ubiquitous Atmel ATMEGA328p microcontroller. When an automobile is detected by the Arduino Fio, a packet will be sent through the onboard XBee Pro radio to the central base station (described in section 2.4). The packet will  contain the direction of the passing vehicle. At the time of detection, a hardware interrupt will also be sent to the license plate recognition system to awaken it from sleep mode.

@@

Once the OpenCV software has detected the license plate, the image will then be cropped as needed until all that remains in the image is the license plate from the car that is entering the lot. That image will then be sent to the central base station via an XBee Pro radio transmission.

\subsection{Central Base Station}
{\color{red}\textbf{NOT COMPLETE}}
The central base station will be the point of contact between the sensor network and the server. The data gathered by the Arduino Fio and Raspberry Pi sensors will be routed to the base station via XBee Pro radios. The base station, which will itself be a Raspberry Pi, will forward the data to the ACMx server using HTTP requests. Forwarding data is the main function of the base station. The base station will run the Raspbian operating system \cite{raspbian}, a lightweight version of Linux. Raspbian is optimized for the Raspberry Pi, so writing shell scripts that will take care of forwarding the data via HTTP to the server will be an easily achievable task. A second function of the base station will be to take periodic overhead photos of the parking lots to help reconcile the sensed data with the actual state of the lots. System administrators will be able to look at the photos to see if the number of cars reported in the lot by the sensors is correct, and fix any errors. See section 2.6 for more detailed information about the error correction functionality of the web application.

@@

The central base station, as described in section 2.4, will be placed in an upper window of the CTLM building, allowing it to take a clear overhead picture of at least one of the two parking lots. For the first deployment we may only have one overhead camera to start off with, although ideally we would have two - one for each lot. Several potential locations exist which allow for a clear line-of-sight to all automobile detection subsystems, one of which is detailed in Figure \ref{fig:deployment}. The location of the base station must put it in range of the XBee Pros mounted on each sensor node, to allow data to be transferred from the detection subsystem to the base station. The locations being considered must also provide internet access to the base station Raspberry Pi through ethernet ports.

>>>>

\section{Related Work {\color{red}\textbf{NOT COMPLETE}} }
The ACMx group at CSM has designed a system named SmartLots to track the number of vehicles in a parking lot. Their system is a wireless sensor network consisting of sensor nodes and a central base station. The sensor nodes are based on the Arduino Fio microcontroller platform equipped with triaxial magnetometers and XBee Pro radios. The base station is a Raspberry Pi connected to the internet via Ethernet. The information collected by the sensor nodes is transmitted to the base station and forwarded to a server running a small web application. The system is described in \cite{stillwell2013} and \cite{parkingWiki}. \cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using the Arduino Fio and the Raspberry Pi. This system uses the IEEE 802.15.4 communications standard to communicate automobile detection data from the sensor nodes to the base station. \cite{parkingWiki} describes the ongoing status of the ACMx project. At the time of writing, the ACMx team has created a website that will display the status of the CTLM upper and lower lots once the monitoring system has been deployed. \cite{parkingWiki} is updated periodically with new information regarding project progress. The project described in this paper augments and extends the ACMx project while collaborating with Stillwell (ACMx president and coauthor of \cite{stillwell2013}).

\section{Project Overview and Methods {\color{red}\textbf{NOT COMPLETE}} }
\begin{figure}
\begin{center}
\includegraphics[width=4.5in]{architecture}
\end{center}
\caption{Overall System Architecture}
\label{fig:system}
\end{figure}
This project is partially an extension of the ACMx SmartLots project. The system consists of several components, each of which will is either an extension of a current SmartLots system or an entirely new subsystem that has been integrated into the existing system. Figure \ref{fig:system} shows the overall architecture of our system. The system relies on three main components:
\begin{enumerate}
\item Detection Subsystem,
\item Central Base Station, and
\item Server.
\end{enumerate}

The detection subsystem consists of a small number of sensor nodes, placed strategically at parking lot entrances. Each sensor node will contain an Arduino Fio equipped with a triaxial magnetometer that will be used to detect when a car passes the sensor. The work to determine the hardware configuration and software for the Arduino Fios has already been done by the ACMx group in \cite{stillwell2013}. Each sensing node is also equipped with a Raspberry Pi single-board computer and a 5 megapixel camera, which take photographs of the license plates of cars as they pass by. When a car passes a sensor node, the magnetometer will detect the car, and the Arduino Fio will send a hardware interrupt to the Raspberry Pi, running a Python script so that the camera can take a picture of the license plate. These images are then run through an OpenCV Python program to reduce the size of the image, through downscaling, cropping, and compression algorithms.

A Raspberry Pi operates as a central base station that receives automobile detection and status data from every sensor node, via transmission from the XBee radios attached to the sensor nodes. The base station Raspberry Pi will be centrally located in a nearby building with internet connectivity. The base station utilizes its network connection to forward the data to the ACMx server.

The server runs a PHP web application with an RESTful HTTP interface that both receives the data from the base station and routes it to the correct part of the application (database, image processor, etc.). This interface will be the method of communication between the server and the base station. When data is received from the base station, the application processes it, stores it in a MySQL database, and displays it appropriately on a user-facing website. An administrator website also exists that provides direct access to license plate data and car counts.

{\color{red}\textbf{NOT COMPLETE}} The following sections describe the architecture and function of each of the aforementioned subsystems in detail. The detection subsystem is broken up into two parts: Automobile Detection (section 2.2) and License Plate Image Acquisition (section 2.3). The server subsystem is also broken into two parts: Server Processing (section 2.5) and Web Application (section 2.6). Lastly, section 2.4 discusses the Central Base Station.

\subsection{Automobile Detection}
{\color{red}\textbf{NOT COMPLETE}}
The automobile detection subsystem will provide a means for detecting ingress and egress of vehicles from a parking lot. This subsystem is to be placed on the side of the road next to each parking lot entrance and exit. Automobiles will be detected using a magnetometer which perceives the induced change in the local magnetic field as the metallic structure of the vehicle passes by, as described in \cite{stillwell2013}. The automobile detection subsystem will be based around the commercially available Arduino Fio 16-bit  platform which utilizes the ubiquitous Atmel ATMEGA328p microcontroller. When an automobile is detected by the Arduino Fio, a packet will be sent through the onboard XBee radio to the central base station (described in section 2.4). The packet will  contain the direction of the passing vehicle. At the time of detection, a hardware interrupt will also be sent to the license plate recognition system to awaken it from sleep mode.

@@

Once the OpenCV software has detected the license plate, the image will then be cropped as needed until all that remains in the image is the license plate from the car that is entering the lot. That image will then be sent to the central base station via an XBee radio transmission.

\subsection{Central Base Station}
{\color{red}\textbf{NOT COMPLETE}}
The central base station will be the point of contact between the sensor network and the server. The data gathered by the Arduino Fio and Raspberry Pi sensors will be routed to the base station via XBee radios. The base station, which will itself be a Raspberry Pi, will forward the data to the ACMx server using HTTP requests. Forwarding data is the main function of the base station. The base station will run the Raspbian operating system \cite{raspbian}, a lightweight version of Linux. Raspbian is optimized for the Raspberry Pi, so writing shell scripts that will take care of forwarding the data via HTTP to the server will be an easily achievable task. A second function of the base station will be to take periodic overhead photos of the parking lots to help reconcile the sensed data with the actual state of the lots. System administrators will be able to look at the photos to see if the number of cars reported in the lot by the sensors is correct, and fix any errors. See section 2.6 for more detailed information about the error correction functionality of the web application.

@@

The central base station, as described in section 2.4, will be placed in an upper window of the CTLM building, allowing it to take a clear overhead picture of at least one of the two parking lots. For the first deployment we may only have one overhead camera to start off with, although ideally we would have two - one for each lot. Several potential locations exist which allow for a clear line-of-sight to all automobile detection subsystems, one of which is detailed in Figure \ref{fig:deployment}. The location of the base station must put it in range of the XBees mounted on each sensor node, to allow data to be transferred from the detection subsystem to the base station. The locations being considered must also provide internet access to the base station Raspberry Pi through ethernet ports.
